---
title: "Reproducible Research: Peer Assessment 1"
author: "William Scott Cochran"
date: "10/17/2014"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r} 
setwd("~/Dropbox/Coursera/Data Science/Reproducible Research/RepData_PeerAssessment1")

# Unzip the data if its not already here
if(! file.exists('activity.csv')) {
        unzip('activity.zip')
}

df <- read.csv('activity.csv', 
               header = TRUE, 
               colClasses = c(steps="numeric",
                              date="Date",
                              interval="numeric"))
library(data.table)
Activity <- as.data.table(df)

stepsPerDay <- aggregate(steps ~ date, Activity, sum)
```

## What is mean total number of steps taken per day?
```{r Histogram}
hist(stepsPerDay$steps, 
     xlab = "Steps Per Day", 
     main ="Histogram of Steps Taken Each Day", 
     col = "red",
     breaks=10)

meanStepsPerDay <- mean(stepsPerDay$steps)
medianStepsPerDay <- median(stepsPerDay$steps)
```

The mean steps taken per day is: `r meanStepsPerDay`

The median steps taken per day is: `r medianStepsPerDay`

## What is the average daily activity pattern?
```{r}
stepsPerInterval <- aggregate(steps ~ interval, Activity, mean)
plot(steps ~ interval, 
     data = stepsPerInterval, 
     type = 'l', 
     xlab="5 Minute Interval", 
     ylab="Mean Number of Steps",
     main="Average Daily Activity")

maxAvgStepsInterval <- 
        stepsPerInterval[which.max(stepsPerInterval$steps),]$interval
maxAvgSteps <- stepsPerInterval[which.max(stepsPerInterval$steps),]$steps
```
Interval number `r maxAvgStepsInterval` has the maximum average number(`r maxAvgSteps`) of steps.


## Imputing missing values
```{r}
missingData <- sum(!complete.cases(Activity))

ActivityImputed <- Activity[,steps := ifelse(is.na(steps), 
                                          mean(steps, na.rm=TRUE),
                                          steps),
                         by=interval]

stepsPerDayImputed <- aggregate(steps ~ date, ActivityImputed, sum)

hist(stepsPerDayImputed$steps, 
     xlab = "Steps Per Day(Imputed NA's)", 
     main ="Histogram of Steps Taken Each Day with Imputed NAs", 
     col = "red",
     breaks=10)

meanStepsPerDayImputed <- mean(stepsPerDayImputed$steps)
medianStepsPerDayImputed <- median(stepsPerDayImputed$steps)


```
There were `r missingData` missing values which were imputed as the mean for their given interval.

The process of imputing mean values for the missing step data coerced the step counts to floats, causing the value of the median steps to be rounded up 1 step.

The Imputed NA mean steps taken per day is: `r meanStepsPerDayImputed`

The Imputed NA median steps taken per day is: `r medianStepsPerDayImputed`


## Are there differences in activity patterns between weekdays and weekends?
```{r}

ActivityImputed[, daytype:= ifelse(weekdays(date)=="Saturday" 
                                   | weekdays(date)=="Sunday",
                                   "weekend", "weekday")]

ActivityImputed[, daytype := as.factor(daytype)]

stepsPerIntervalImputed <- aggregate(steps ~ interval + daytype, ActivityImputed, mean)

library(lattice)
xyplot(steps ~ interval | daytype, stepsPerIntervalImputed, type='l')
```
